---
title: "DSC241 HW1"
author:
  - Zihe Liu
  - A59026217
output: 
  pdf_document:
    latex_engine: xelatex
date: "2024-01-19"
---


\includegraphics[angle=270]{/D/hw1.jpg}

# Write a function confBand(x,y,conf=0.95) that takes as inputs a predictor vector x = (x1, . . . , xn) and a response vector y = (y1, . . . , yn), and returns a plot with the points (x1, y1), . . . ,(xn, yn), the least squares line, and two confidence bands: pointwise and simultaneous, both at their corresponding level conf.

# Apply your function to the regression of mpg on hp in the 04cars dataset and explain the result.
```{r}
# load libraries
library("ggplot2")
# library("MASS")
# library(dplyr) 
```

```{r}
# Function to plot the points (x1, y1), . . . ,(xn, yn), the least squares line, and two confidence bands: pointwise and simultaneous
confBand <- function(x, y, conf = 0.95) {
  
  # compute linear model 
  Model <- as.formula(y ~ x)
  fit1 <- lm(Model) 
  
  # compute design matrix 
  X1 <- model.matrix(fit1) 
  
  # general linear hypotheses 
  cht1 <- multcomp::glht(fit1, linfct = X1)  
  
  # simultaneous confidence bands 
  data <- data.frame(x, y)
  # sim_CB <- simultaneous_CBs(fit1, data)
  
  l = length(x)
  alpha = 1 - conf
  alpha1 = alpha/l
  # Plot setosa data with simultaneous confidence bands
  ggplot(data = data, aes(x = x, y = y)) +
    geom_point() +
    geom_smooth(aes(color = "Least Squares Line"), method = "lm", formula = y ~ x, level = conf) +
    geom_smooth(aes(fill = "Simultaneous Confidence Bands"), method = "lm", formula = y ~ x, level = 1 - alpha1) +
    labs(x = "hp", y = "mpg", title = "Linear Model with Confidence Bands") +
    ggtitle("04cars Dataset in the Least Squares Line with Pointwise and Simultaneous Confidence Bands") +
    scale_color_manual(values = "black", name = "Legend", labels = "Pointwise Confidence Bands") +
    scale_fill_manual(values = "lightgray", name = "Legend", labels = "Simultaneous Confidence Bands")
}
```

```{r}
#Load '04cars' dataset
#tmp = as.data.frame(tmp)
#names(tmp) = c("hp","mpg","wt","len","wd") # abbreviate names
#file_path <- "D:/Liuuu/Graduate@UCSD/Courses/winter/DSC_241_Statistical Models/hw/04cars.rda"
#file_path <- "/D/04cars.rda"

load("/D/04cars.rda")
tmp = dat[,c(13,15)] # extract selected variables
tmp = tmp[complete.cases(tmp),] # extracts complete cases
tmp = as.data.frame(tmp)
names(tmp) = c("hp","mpg") # abbreviate names

# mpg <- get("mpg", envir = .GlobalEnv)
#tmp$mpg
#hp <- get("hp", envir = .GlobalEnv)

# Assuming hp and mpg are vectors
#data <- data.frame(hp, mpg)
confBand(tmp$hp, tmp$mpg)
```

# There is a ngative linear relationship between hp and mpg.

# Let n = 100 and draw x1, . . . , xn iid from Unif(0,1), which stay fixed in what follows. Repeat the following experiment N = 1000 times:
# (i) Generate yi = 3 + 0.5xi + ϵi , i = 1, . . . , n, with ϵ1, . . . , ϵn iid N (0, 0.2).
# (ii) Compute the 95% pointwise and simultaneous confidence bands and record whether they contain the true line or not.
# Summarize the result of this numerical experiment by returning the proportion of times (out of N) that the pointwise and simultaneous confidence bands contained the true line. Explain the results.

```{r}
# Function to generate data and compute confidence bands
simulate_experiment <- function(n, N, conf) {
  contains_true_line_pointwise <- contains_true_line_simultaneous <- rep(0, N)

  for (i in 1:N) {
    # Generate xi
    x <- runif(n, min = 0, max = 1)

    # Generate epsilon_i
    epsilon <- rnorm(n, mean = 0, sd = 0.2)

    # Generate yi
    y <- 3 + 0.5 * x + epsilon

    # Fit linear model
    linear_model <- lm(y ~ x)

    # Compute pointwise confidence bands
    ci_pointwise <- predict(linear_model, interval = "confidence", level = conf)
   

    # Compute simultaneous confidence bands
    l = length(x)
    alpha = 1 - conf
    alpha1 = alpha/l
    sim_CB <- predict(linear_model, interval = "confidence", level = 1-alpha1)


    # Check if the true line (y = 3 + 0.5x) is within the bands
    contains_true_line_pointwise[i] <- all(ci_pointwise[, "upr"] >= 3 + 0.5 * x & ci_pointwise[, "lwr"] <= 3 + 0.5 * x)
    contains_true_line_simultaneous[i] <- all(sim_CB[, "lwr"] <= 3 + 0.5 * x & sim_CB[, "upr"] >= 3 + 0.5 * x)
  }

  return(list(
    contains_true_line_pointwise = contains_true_line_pointwise,
    contains_true_line_simultaneous = contains_true_line_simultaneous
  ))
}
```

```{r}

# Set parameters
n <- 100
N <- 1000
conf <- 0.95

# Run simulation
results <- simulate_experiment(n, N, conf)

# Calculate proportions of containing the true line
prop_pointwise <- mean(results$contains_true_line_pointwise)
prop_simultaneous <- mean(results$contains_true_line_simultaneous)

# Display results
cat("Proportion of pointwise confidence bands containing the true line:", prop_pointwise, "\n")
cat("Proportion of simultaneous confidence bands containing the true line:", prop_simultaneous, "\n")
```
# Because simultaneous confidence bands are larger than pointwise confidence bands therefore the proportion of simultaneous confidence bands containing the true line is bigger.
