---
title: "DSC241 HW3"
author:
- Zihe Liu
- A59026217
date: "2024-01-31"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---


# 1.(1) Compute the condition number of the canonical design matrix. Produce a useful plot for visualizing the result of these computations. Offer brief comments.
```{r}

condition_number_calculation <- function(n,p,raw){
  
  x = runif(n)
  X = poly(x, degree = p, raw = raw)
  X = cbind(rep(1, n), X) 
  d = svd(X)$d 
  con = max(d)/min(d)
  return (con)

}

raw <- TRUE
maxP <- 20
P <- seq(1, maxP)
N <- c(30, 50, 100, 200, 500, 1000)

for (i in 1:length(N)){
  cons <- numeric(maxP)
  for (j in 1:maxP){
    cons[j] <- condition_number_calculation(N[i],P[j],raw)
  }
  plot(P, cons, main = paste("n=", N[i]), xlab = "polynomial of degree p", ylab = "condition number")
}

```

## Comments: When p increases and n remains unchanged, condition number of the design matrix increases.

## A higher condition number implies a matrix that is closer to singularity, and consequently, numerical computations may become less stable.

# 1.(2) What is the condition number of X when using orthogonal polynomials instead?

```{r}
raw <- FALSE
con_p <- numeric(length(N))
for (i in 1:length(N)){
  cons <- numeric(20)
  for (j in 1:20){
    cons[j] <- condition_number_calculation(N[i],P[j],raw)
    print(cons[j])
  }
  con_p[i] <- cons[j]
  plot(P, cons, main = paste("n=", N[i]), xlab = "polynomial of degree p", ylab = "condition number")
}

plot(N, con_p, xlab = "N", ylab = "condition number")
```

## Comments: When n remains unchanged, condition number doesn't change as p increase. When N increases, condition number increases.

# 2.(1) piecewiseConstant function
```{r}
piecewiseConstant <- function(x, y, L, plot = TRUE){
  K = quantile(x, seq(0, 1, len = 2^L), type=1)
  len = 2^L - 1
  pts = rep(0,len*2)
  val = rep(0,len*2)
  for (j in 1:len){
    I = (K[j] < x)&(x <= K[j+1])
    fit = lm(y[I] ~ 1)
    pts[2*j-1] = K[j]
    pts[2*j] = K[j+1]
    val[2*j-1] = coef(fit)
    val[2*j] = coef(fit)
  }
  fit0 <- list(pts = pts, val = val)

  if (plot){
    if (L == 2){
      lines(fit0$pts, fit0$val, col="blue", lwd = 3)
    }
    
    if (L == 3){
      lines(fit0$pts, fit0$val, col="green", lwd = 3)
    }
    
    if (L == 4){
      lines(fit0$pts, fit0$val, col="red", lwd = 3)
    }
  }
    
}
```
# 2.(2) Apply the function to explaining City Mpg as a piecewise constant function of Horsepower in the 04cars dataset. 
```{r}
## get the data and plot scatter plot
load("/D/04cars.rda")
tmp = dat[,c(13,14)] # "Horsepower","City_MPG"
tmp = tmp[complete.cases(tmp),]
tmp = as.data.frame(tmp)

x = tmp$Horsepower
y = tmp$City_MPG

par(mfrow = c(1, 1))
plot(x, y, main = "Piecewise constant fit", xlab = "Horsepower", ylab = "City MPG")
for (i in 2:4){
  piecewiseConstant(x, y, i, plot = TRUE)
}
legend("topright", legend = c("L=2", "L=3", "L=4"), col = c("blue", "green", "red"), lwd = 3)
```